---
title: "Rural Innovation Network"
favicon: "./assets/favicon.ico"
format: html
categories:
  - R Package
  - Network Data
  - Rural Development
package:
  name: cori.data.rin
  githubUrl: https://github.com/ruralinnovation/cori.data.rin
  installCommand: 'devtools::install_github("ruralinnovation/cori.data.rin")'
  status: stable
  version: 1.0.1
  maintainer: John Hall
  featured: false
permalink: /packages/cori-data-rin/
resources:
  - "./assets"
  - "./favicon.ico"
  - "./frontend/dist"
tags:
  - RIN
  - Rural Innovation Network
  - Community Data
  - Service Areas
  - County-Level
  - Targets Pipeline
  - S3
usesDatasets:
  - rin-communities
usesPackages: []
usesResources: []
include-after-body:
  - file: "./frontend/dist/index.html"
#   text: |
#     <script>
#       alert('This is triggered by a custom script tag inserted after </body>');
#     </script>
# include-before-body:
#   text: |
#     <script crossorigin src="https://unpkg.com/react@18/umd/react.production.min.js"></script>
#     <script crossorigin src="https://unpkg.com/react-dom@18/umd/react-dom.production.min.js"></script>
# server: shiny
# ERROR: index.qmd uses server: shiny so cannot be included in a website project (shiny documents require a backend server and so can't be published as static web content).

---

<div class="embedded-content side-bar">

<!--
  -- And here's some more stuff to say about the network.
  -- Should we drop a micro dashboard here?
  -->

```{r}
#| eval: true
#| echo: false
#| message: false
#| output: true
#| results: asis

tableau_client_id <- Sys.getenv("TABLEAU_CLIENT_ID")
tableau_secret_id <- Sys.getenv("TABLEAU_SECRET_ID")
tableau_secret_value <- Sys.getenv("TABLEAU_SECRET_VALUE")
tableau_url <- Sys.getenv("TABLEAU_URL")
tableau_user <- Sys.getenv("TABLEAU_USER")

cat(paste0('
  <tableau-viz
    src="', tableau_url, '/views/impact-dashboard/NetworkMaturity2?:origin=card_share_link&:embed=n4?:origin=card_share_link"
    device="desktop"
    width="640px"
    height="calc(100vh - 190px)"
    style="margin: 0 -2px; background-color: transparent !important;"
    toolbar="bottom"
    tableau_client_id="', tableau_client_id, '"
    tableau_secret_id="', tableau_secret_id, '"
    tableau_secret_value="', tableau_secret_value, '"
    tableau_user="', tableau_user, '">
  </tableau-viz>
  <script type="application/javascript">
  setTimeout(() => { document.querySelector("main#quarto-document-content").style.overflow = "hidden" }, 533)
  </script>
'
))
```

<!--
```r
#| eval: false
sliderInput("bins", "Number of bins:",
            min = 1, max = 50, value = 30)
plotOutput("distPlot")
```

```r
#| context: server
#| eval: false
output$distPlot <- renderPlot({
  x <- faithful[, 2]  # Old Faithful Geyser data
  bins <- seq(min(x), max(x), length.out = input$bins + 1)
  hist(x, breaks = bins, col = 'darkgray', border = 'white')
})
```
-->

</div>

## Overview
 
The `cori.data.rin` package exports `rin_service_areas` — the canonical
mapping from Rural Innovation Network (RIN) community names to their
associated U.S. counties, keyed by 5-digit FIPS code. This is the
authoritative county crosswalk used across CORI tools whenever RIN
geography needs to be joined to county-level data.
 
::: callout-note
## Why Use This Package
 
-   **Canonical RIN geography** — single source of truth for which
    counties belong to each RIN community
-   **FIPS-ready** — `geoid_co` is a standard 5-digit county FIPS code
    for direct joins to ACS, BDS, COG, or any other county-level dataset
-   **Primary county flag** — distinguishes a community's primary county
    from secondary service area counties
-   **Always current** — MDA refreshes the dataset from the
    Comms-maintained Google Sheet whenever community boundaries change
:::
 
## Installation
 
```{r}
#| eval: false
 
devtools::install_github("ruralinnovation/cori.data.rin")
```
 
## Quick Start
 
```{r}
#| eval: false
 
library(cori.data.rin)
library(dplyr)
 
# Load the county-level service area crosswalk
rin_service_areas
 
# How many counties does each RIN community serve?
rin_service_areas |>
  count(rin_community, name = "n_counties") |>
  arrange(desc(n_counties))
 
# Primary counties only
rin_service_areas |>
  filter(primary_county_flag == "Yes")
 
# Join RIN community membership onto another county dataset
county_data |>
  left_join(rin_service_areas, by = c("geoid" = "geoid_co"))
```
 
## Data Structure
 
### `rin_service_areas`
 
A data frame with one row per community–county pair:
 
| Column | Type | Description |
|---|---|---|
| `geoid_co` | character | County FIPS (5-digit: STATEFP + COUNTYFP) |
| `rin_community` | character | CORI name for the RIN community |
| `county` | character | County name |
| `primary_county_flag` | character | `"Yes"` / `"No"` — primary county for this community |
| `data_run_date` | Date | Date the dataset was last refreshed from the source sheet |
 
## Data Coverage
 
| Dimension | Coverage |
|---|---|
| **Geography** | County (5-digit FIPS) |
| **Unit** | One row per RIN community × county pair |
| **Source** | Comms-maintained Google Sheet, updated as communities join or expand |
| **Update Frequency** | On demand — MDA runs `targets::tar_make()` after Comms updates the sheet |
 
## Use Cases
 
-   **County-level joins** — Attach ACS, BDS, COG, or any other
    county-level dataset to RIN communities via `geoid_co`
-   **Network coverage analysis** — Identify which counties are served
    by one or more RIN communities, and which communities share counties
-   **Reporting and mapping** — Provide the spatial layer for RIN
    community dashboards and the [RIN map frontend](https://github.com/ruralinnovation/rin-map)
-   **Agent knowledge base** — Supplies the geographic context layer for
    CORI's AI agent to answer questions about RIN community locations
 
## Data Refresh (Admin)
 
The dataset is generated by a `targets` pipeline maintained by MDA.
After Comms updates the [source Google Sheet](https://docs.google.com/spreadsheets/d/1fDGclR7GqncaLdkEipOJLLwDn3h4gYHVRcJtP_K1wcU):
 
```{r}
#| eval: false
 
# From the cori.data.rin project root
targets::tar_make()
```
 
This rebuilds `rin_service_areas`, writes `data/rin_map.json`, and
uploads outputs to S3 (`cori-risi-apps` for the RIN map frontend,
`cori.agent.kb` for the agent knowledge base). Once verified on dev,
the updated files are copied to the production S3 prefix.
 
## Related Resources
 
### Datasets
 
-   RIN Communities *(slug: `rin-communities` — node pending)* —
    Community and service area data maintained by Comms in Monday.com
    and Google Sheets
 
### External Links
 
-   [RIN Map Repository](https://github.com/ruralinnovation/rin-map)
-   [Monday.com Board](https://ruralinnovation-group.monday.com/boards/6951894369)
 
::: callout-note
## Dangling references
 
The following slugs are referenced by this package but do not yet have
nodes in Dataverse. They are intentionally preserved as future content
needs:
 
-   `dataset/rin-communities`
:::
 
---
 
**Contact:** John Hall (john.hall@ruralinnovation.us)
 
For bug reports or feature requests, please open an issue on
[GitHub](https://github.com/ruralinnovation/cori.data.rin/issues).
